---
title: "Automatic report"
subtitle: "Escola Nacional de Saúde Pública"
author: "A. Peralta-Santos"
date: today
date-format: long
lang: pt 
format: 
  docx: 
    reference-doc: templates/template1.docx
execute: 
  echo: false
  warning: false
project:
  output-dir: outputs/documents
bibliography: references.bib
csl: templates/vancouver.csl
---



```{r}
rm(list = ls(all.names = TRUE)) # limpa todos os objetos, incluindo os ocultos
```

```{r}
#install.packages("broom")
#install.packages("MLDataR")
#install.packages("gtsummary")
# check the website https://www.danieldsjoberg.com/gtsummary/ 
#install.packages("report")

library(tidyverse)
library(broom)
library(MLDataR)
library(janitor)
library(gtsummary)
library(gt)
library(report)
```

```{r}
hd <- MLDataR::heartdisease
hd<- clean_names(hd)
```

{{< pagebreak >}}

# Introduction {.unnumbered}

This is a smart report generated with R on the date `r format(Sys.time(), '%d %B %Y')`. We used a dataset in the MLDataR called Heart Disease. In this report we will combine what we have learnt so far in the R course.



{{< pagebreak >}}

# Methods {.unnumbered}

## Study Design and Data Source {.unnumbered}

This study employs a retrospective analysis design using the Heart Failure Dataset from MLDataR. This dataset is an open-source, publicly available resource that compiles data from heart failure patients, offering comprehensive insights into clinical, demographic, and laboratory features that may be associated with heart failure outcomes.

## Data Collection

The Heart Failure Dataset from MLDataR was downloaded and parsed into a format amenable to our analysis pipeline.

We created a logistic regression model, with the following equation

$$ E(Y_i) = \beta_0 + \beta_{1}  AGE_i + \beta_{2}  SEX_i +   \beta_{3}  BP_i +   \beta_{4}  CHOL_i + \epsilon_i$$

{{< pagebreak >}}

# Results {.unnumbered}

```{r}
hd_sum <- hd %>%
  summarise(
    mean_age=mean(age)
  )

age_sum <- round(hd_sum$mean_age,1)

```

```{r}
hd_sim1 <- hd %>%
  summarise(
    sum__cases=n()
  )
hd_sim1 <- round(hd_sim1$sum__cases,1)
```

```{r}
hd<- hd %>%
  mutate(heart_disease_yesno = ifelse(heart_disease==1, "Yes", "No"))
```

## Descriptive Statistics {.unnumbered}

Our data set as `r hd_sim1` patients with a mean age of `r age_sum`


```{r}
table2 <-
  tbl_summary(
    hd,
    include = c(sex, age,  resting_ecg, resting_bp, cholesterol),
    by = heart_disease_yesno, # split table by group
    missing = "no" # don't list missing data separately
  ) %>%
  add_overall() %>% 
  modify_spanning_header(c("stat_1", "stat_2") ~ "**Heart Disease**") %>%
  add_n() %>% # add column with total number of non-missing observations
  add_p() %>% # test for a difference between groups
  modify_header(label = "**Variable**") %>% # update the column header
  modify_caption("** Patient Characteristics **") %>%
  bold_labels()

table2
```

{{< pagebreak >}}

## Model {.unnumbered}

```{r}
model3_logit <- glm(
  heart_disease ~ age + as.factor(sex)+ relevel(factor(resting_ecg), ref = "Normal") + resting_bp+ cholesterol, 
  family="binomial", 
  data=hd
)

model3_tidy <- tidy(model3_logit, exponentiate = TRUE, conf.int = TRUE)

a<-report(model3_logit)
```

`r a`

```{r}
model3_tidy <- model3_tidy %>%
  select(term, estimate, p.value, conf.low, conf.high) %>%
    mutate(
    term = case_when(
      term =="(Intercept)" ~ "Intercept",
      term =="age" ~ "Age",
      term =="as.factor(sex)M" ~ "Male",
      term =="resting_bp" ~ "Blood preassure",
      term =='relevel(factor(resting_ecg), ref = "Normal")ST' ~ "ST ECG, ref. Normal",
      term =='relevel(factor(resting_ecg), ref = "Normal")LVH' ~ "LVH ECG, ref. Normal",
      term =="cholesterol" ~ "Cholesterol",
    )
  ) %>% rename(
    "Risk factor"=term,
    "Odds ratio"=estimate,
    "p value" = p.value,
    "Low 95%CI"=conf.low,
    "High 95%CI"=conf.high
  ) 
```

{{< pagebreak >}}

The plot bellow is the relationship between age and resting blood pressure by heart disease

```{r}
p1<- ggplot(
  data = hd, 
  mapping = aes(
    x=age, 
    y=cholesterol, 
    colour=as.factor(sex))
  )+ 
  geom_point(alpha=0.5) +
  geom_smooth(method = loess, se= FALSE) +
  scale_colour_discrete(labels=c('Female', 'Male')) +
    labs(
      title = "Age and Cholesterol",
      subtitle = "By patient sex",
      y = "Cholesterol",
      x = "Age", 
      caption = "Source: ML datasets",
      colour=""
    ) +
   theme_minimal() +
  theme(legend.position="bottom")

p1
```

{{< pagebreak >}}

```{r}
# Make a display table with the `islands_tbl`
# table; put a heading just above the column labels
gt_model3_tidy <- 
  gt(model3_tidy) %>% 
  tab_header(
    title = md("**Odds Ratio for Heart Failure**"),
    subtitle = md("*Based on a binomial GLM adjusted for confounding*")
  ) %>% 
  opt_align_table_header("left") %>% 
  fmt_number(
    decimals = 3
  ) %>% opt_row_striping() %>% 
  cols_width(
    everything() ~ px(150)
  ) %>% 
  tab_source_note(
    source_note = "Source: MLDataR dataset"
  ) %>% 
  tab_footnote(
    footnote = "ECG was performed with a resting patients",
    locations = cells_body(columns = "Risk factor", rows = 4:5)
  )

gt_model3_tidy
```
{{< pagebreak >}}

# Discussion  {.unnumbered}

This study utilizes a logistic model to predict heart disease using multiple factors such as age, sex, resting electrocardiogram (ECG) results, resting blood pressure (BP), and cholesterol levels. The logistic model is a form of regression analysis used when the dependent variable is a binary outcome - in this case, the presence or absence of heart disease.

The logistic model has a moderate explanatory power with a Tjur's R^2 of 0.20. This means that approximately 20% of the variability in the presence of heart disease can be explained by the variables included in this model.

The model's intercept is at -4.46 with a 95% Confidence Interval (CI) ranging from -5.80 to -3.15. This indicates that at the baseline level (i.e., when age, sex, resting ECG, resting BP, and cholesterol are zero), the log-odds of having heart disease is -4.46. This value is statistically significant, with a p-value less than 0.001.

Among the predictors, age and sex are statistically significant and positively associated with heart disease. The positive beta values of 0.06 and 1.47 respectively suggest that with each additional year of age, and being male, the log-odds of heart disease increase [@cowin2005cadherins]. 

Resting ECG, when relevelled to "Normal", shows two categories LVH (Left Ventricular Hypertrophy) and ST (ST wave abnormality) are not statistically significant in predicting heart disease. This means the odds of heart disease do not significantly change with these specific ECG results as compared to a normal ECG result [@litaker2005context].

Resting BP is statistically significant and positively associated with heart disease. The positive beta value of 0.00848 suggests that for each unit increase in resting BP, the log-odds of having heart disease slightly increase.

It's important to note that these results should be interpreted in the context of the data used, and any clinical implications should be considered cautiously. Other factors not included in the model may also be significant predictors of heart disease, and the relationship between cholesterol and heart disease may need further investigation to understand why it is negatively associated in this model. Additionally, the logistic model, like any other model, is a simplification of reality and should be used as one piece in a larger puzzle when making predictions about heart disease.

{{< pagebreak >}}

# References 

::: {#refs}
:::
