---
title: "Day 2 | Smart reports | HTML"
author: "A. Peralta-Santos"
date: today
date-format: long
format:
  html:
    smooth-scroll: true
    self-contained: true
    toc: true
    toc-location: left
theme: cosmo
smooth-scroll: true
execute:
  warning: false
  echo: false
project:
  output-dir: outputs/documents
---



```{r}
rm(list = ls(all.names = TRUE)) # limpa todos os objetos, incluindo os ocultos
```

```{r}
#install.packages("table1")
#install.packages("openxlsx")
#install.packages("flextable")

library(tidyverse)
library(broom)
library(MLDataR)
library(table1)
library(openxlsx)
library(here)
library(flextable)

```

```{r}
hd <- MLDataR::heartdisease

hd<- janitor::clean_names(hd)
```

# Introduction

This is a smart report generated with R on the date `r format(Sys.time(), '%d %B %Y')`. We used a dataset in the MLDataR called Heart Disease. In this report we will combite what we have learnt so far in the R course.


\newpage

# Methods

We created a logistic regression model $x=y$

$$ E(Y_i) = \beta_0 + \beta_{1}  AGE_i + \beta_{2}  SEX_i +   \beta_{3}  BP_i +   \beta_{4}  CHOL_i + \epsilon_i$$

\newpage

# Results

```{r}
hd_sum <- hd %>%
  summarise(
    mean_age=mean(age)
  )

age_sum <- round(hd_sum$mean_age,1)
```

```{r}
hd_sim1 <- hd %>%
  summarise(
    sum__cases=n()
  )
hd_sim1 <- round(hd_sim1$sum__cases,1)
```

## Descriptive Statistics

Our data set as `r hd_sim1` patients with a mean age of `r age_sum`. 

teste por sugestao do tiago `r round(hd_sum$mean_age,1)` 

```{r}
t1 <- table1(
  ~ factor(sex) + age + factor(resting_ecg) + cholesterol + resting_bp | heart_disease, 
  data = hd,
  # footnote = "Source: MLData R",
  caption = "Descriptive variables by heart disease"
)



table1 <- as.data.frame(t1)

write.xlsx(
  table1,
  file = here("output", "datasets", paste0("table1_", Sys.Date(),".xlsx")) 
)
```

```{r}
table1(
  ~ factor(sex) + age + factor(resting_ecg) + cholesterol + resting_bp | heart_disease, 
  data=hd,
  #footnote = "Source: MLData R",
  caption = "Descriptive variables by health disease"
)
```

\newpage

## Model

```{r}
model3_logit <- glm(
  heart_disease ~ age + as.factor(sex) + relevel(factor(resting_ecg), ref = "Normal") + resting_bp + cholesterol, 
  family="binomial",
  data = hd
)

model3_tidy <- tidy(model3_logit, exponentiate = TRUE, conf.int = TRUE)
```

```{r}
model3_tidy <- model3_tidy %>%
  select(term, estimate, p.value, conf.low, conf.high) %>%
    mutate(
    term = case_when(
      term =="(Intercept)" ~ "Intercept",
      term =="age" ~ "Age",
      term =="as.factor(sex)M" ~ "Male",
      term =="resting_bp" ~ "Blood preassure",
      term =='relevel(factor(resting_ecg), ref = "Normal")ST' ~ "ST ECG, ref. Normal",
      term =='relevel(factor(resting_ecg), ref = "Normal")LVH' ~ "LVH ECG, ref. Normal",
      term =="cholesterol" ~ "Cholesterol",
    )
  ) %>% rename(
    "Risk factor"=term,
    "Odds ratio"=estimate,
    "p value" = p.value,
    "Low 95%CI"=conf.low,
    "High 95%CI"=conf.high
  ) 
```

The plot bellow is the relationship between age and resting blood pressure by heart disease

```{r}
p1<- ggplot(
  data = hd,
  mapping = aes(
    x=age,
    y=resting_bp,
    colour=as.factor(heart_disease))
  )+ 
  geom_point(alpha=0.5) +
  geom_smooth(
    method = loess,
    se= FALSE
    ) +
  scale_colour_discrete(labels=c('No heart disease', 'Heart disease')) +
  labs(
    title = "Age and resting blood pressure",
    subtitle = "By having Heart Disease",
    y = "Resting bood preassure",
    x = "Age", 
    caption = "Source: ML datasets",
    colour="") +
  theme_minimal() +
  theme(legend.position="bottom")

p1
```

```{r}
set_flextable_defaults(
  font.size = 10, 
  theme_fun = theme_vanilla,
  padding = 6,
  background.color = "#f7f7f7")
```

The table bellow shows the Odds Ratio

```{r}
flextable(model3_tidy) %>%
  colformat_double( digits = 2) %>%
  set_caption("Odds Ratio for Heart Failure") %>%
  set_table_properties(layout = "autofit")
```

## The end
